---
title: "Homework 4"
author: "Mengkun Chen (mengkun21@vt.edu)"
output: pdf_document
---

### Problem 3

```{r message=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(ggplot2)
```

#### Part A

```{r message=FALSE, warning=FALSE}
thick<-fread("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/ThicknessGauge.dat", 
             header = FALSE, data.table = FALSE)
```

- Rename the column: column 2 and 3 are measurements under operator 1, column 4 and 5 are measurements under operator 2, column 6 and 7 are measurements under operator 3. 

- Compute mean thickness for 10 different parts using different operators. 

- Keep the mean thickness as the measure of wall thickness

```{r}
thick<- thick %>%
  rename(Part = V1, G11 = V2, G12 = V3, G21 = V4, G22 = V5, G31 = V6, G32 = V7) %>% 
  mutate(G1 = (G11+G12)/2, G2 = (G21+G22)/2, G3 = (G31+G32)/2) %>% 
  select(Part, G1, G2, G3)

knitr::kable(thick, digits = 4, caption = "Wall Thickness")
```

Summary of dataset: 

```{r}
knitr::kable(summary(thick[,-1]), digits = 4, caption = "Summary of Wall Thickness")
```

Draw bar plot for wall thickness of 10 parts of the wall. 

```{r}
thick %>% gather(key = Guage, value = measure, G1:G3) %>% 
  ggplot(aes(x = Guage, y = measure)) + geom_boxplot() + 
  theme_light()
```

Operator 3 tend to get higher thickness measurements. 


#### Part B

```{r message=FALSE, warning=FALSE}
bbweight<-fread("http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat", header = FALSE, data.table = FALSE)
colnames(bbweight)<-rep(c("Body Wt", "Brain Wt"), 3)
```

- Combine all the body weight and brain weight into two separate columns. 

```{r}
body<-c(bbweight[,1],bbweight[,3],bbweight[,5])
brain<-c(bbweight[,2],bbweight[,4],bbweight[,6])
bb<-data.frame("Body Wt" = body, "Brain Wt" = brain)
knitr::kable(head(bb), digits = 4, caption = "Body-Brain Weight")
```

Summary of dataset: 

```{r}
knitr::kable(summary(bb), digits = 4, caption = "Body-Brain Weight Summary")
```

From the summary table, the dataset may contain several mis-recorded observation, which are far away from the majority observations. 

Draw dot plot to show the relationship between body weight and brain weight ignorning the observations whose body weight and brain weight are greater than the third quartile respectively. 

```{r}
plot(bb$Body.Wt, bb$Brain.Wt, xlab = "Body Weight", ylab = "Brain Weight", 
     main = "Brain Weight vs. Body Weight", xlim = c(0,40), ylim = c(0,130))
lm(bb$Brain.Wt~bb$Body.Wt)
```

Brain weight and body weight are positive related ignorning some obvious outliers. 


#### Part C

```{r message=FALSE, warning=FALSE}
jump<-fread("http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat", data.table = FALSE, header = FALSE)
colnames(jump)<-rep(c("Year", "Long Jump"),4)
```

- Combine all year and long jump columns into two seperate columns. 

```{r}
year<-c(jump[,1], jump[,3], jump[,5], jump[,7])
long.jump<-c(jump[,2], jump[,4], jump[,6], jump[,8])
lj<-data.frame("Year" = year, "Long Jump" = long.jump)
knitr::kable(head(lj), caption = "Gold Medal Performance for Long Jump")
```

Summary of Gold Medal performance for Olympic Men's Long Jump. 

```{r}
knitr::kable(summary(lj), caption = "Summary of Gold Medal Performance for Long Jump")
```

```{r}
plot(lj$Year+1900, lj$Long.Jump, xlab = "Year", ylab = "Long Jump Performance", 
     main = "Gold Medal Performance for Long Jump")
```

Gold Medal performance for Olympic Men's Long Jump increase as year increase. 


#### Part D

```{r message=FALSE, warning=FALSE}
yield<-fread("http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat", 
             header = FALSE)
colnames(yield)<-c("Variety", "D1", "D2", "D3")
```

- Separate tomato yields under different densities. 

```{r message=FALSE, warning=FALSE}
d1<-yield %>% 
  select(D1) %>% 
  separate(col = D1, into = c("y1", "y2", "y3"), sep = ",") %>% 
  mutate(Density = c("D1","D1"))
d2<-yield %>% 
  select(D2) %>% 
  separate(col = D2, into = c("y1", "y2", "y3"), sep = ",") %>% 
  mutate(Density = c("D2","D2"))
d3<-yield %>% 
  select(D3) %>% 
  separate(col = D3, into = c("y1", "y2", "y3"), sep = ",") %>% 
  mutate(Density = c("D3","D3"))
tomato<-cbind("Variety" = rep(c("v1", "v2"), 2), rbind(d1, d2, d3)) %>% 
  gather(key = y, value = yield, y1:y3) %>% 
  select(Variety, Density, yield)
knitr::kable(head(tomato), caption = "Tomator Yield")
```

Summary of tomato yield: 

```{r}
knitr::kable(summary(tomato), caption = "Summary of Tomato Yield")
```

```{r}
tomato %>% ggplot(aes(x = factor(Density), y = as.numeric(yield), 
                      col = factor(Variety))) + 
  geom_point() + 
  theme_classic()
```

Variety 2 has lower yield than variety 1. Tomato yields increase when planting density increase. 


#### Part E

```{r message=FALSE, warning=FALSE}
larvae<-fread("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LarvaeControl.dat", header = TRUE, data.table = FALSE)
```

- Separate data according to age. 

- Rearrange the dataset into four columns: block, treat, count and age. 

```{r}
A1<-larvae[,1:6] %>% 
  rename("T1" = "1", "T2" = "2", "T3" = "3", "T4" = "4", "T5" = "5") %>% 
  gather(key = Treat, value = count, T1:T5) %>% 
  mutate(Age = rep(1, 40))
A2<-larvae[,c(1,7:11)] %>% 
  rename("T1" = "\t1", "T2" = "2", "T3" = "3", "T4" = "4", "T5" = "5") %>% 
  gather(key = Treat, value = count, T1:T5) %>% 
  mutate(Age = rep(2, 40))
larvae_count<-rbind(A1, A2)
knitr::kable(head(larvae_count), caption = "Larvae Counts")
```

Summary of dataset: 

```{r}
larvae_count$Block<-as.factor(larvae_count$Block)
larvae_count$Treat<-as.factor(larvae_count$Treat)
larvae_count$Age<-as.factor(larvae_count$Age)
knitr::kable(summary(larvae_count), caption = "Summary of Larvae Counts")
```

```{r message=FALSE, warning=FALSE}
larvae_count %>% ggplot(aes(x = Treat, y = count, col = Age, alpha = Block)) + 
  geom_point() + 
  theme_classic()
```

Larvae counts are larger in age 2 than those in age 1. Blocks have significant effect on larvae counts and the effect decrease as block number increase. No obvious treatment effect is shown in the plot. 